Forcing for Isabelle

### 2018-12-29 LICS/intro_m.tex
eds varias

The most important achievement of this paper is the
mechanization in the proof assistant \emph{Isabelle} of a proof of the
Axiom of Separation in 
generic extensions by using the ``fundamental'' theorems of forcing.
En route
to this, we also formalized the satisfaction by $M[G]$ of
Extensionality, Foundation, Union, and Infinity. % We have already
% proved Pairing in the first report of our project
% \cite{2018arXiv180705174G}.

@@

The \emph{Formal Abstracts} project~\cite{hales-fabstracts} proposes
the formalization of complex pieces of mathematics by writing the
statements of 
results and the material upon which they are based (definitions,
propositions, lemmas), but ommiting the proofs. In this work we
partially adhere to this vision to delineate our formalization
strategy:
since the proofs that the  axioms hold in generic extensions
is independent of the \emph{proofs} of the fundamental theorems of
forcing, we assumed these results for the time being. Let us remark
that those theorems depend on the definition of a function $\forceisa$
from formulas to formulas which is, by itself, quite demanding; the
formalization of it and of the fundamental theorems of forcing roughly
comprises barely less than a half of our full project.

>>>>

The most important achievement of this paper is the
mechanization in the proof assistant \emph{Isabelle} of a proof of the
Axiom of Separation in 
generic extensions by using the ``fundamental'' theorems of forcing.
En route
to this, we also formalized the satisfaction by $M[G]$ of
Extensionality, Foundation, Union, and Infinity. % We have already
% proved Pairing in the first report of our project
% \cite{2018arXiv180705174G}.
We have  followed \cite{kunen2011set} for this purpose.

@@

The \emph{Formal Abstracts} project~\cite{hales-fabstracts} proposes
the formalization of complex pieces of mathematics by writing the
statements of 
results and the material upon which they are based (definitions,
propositions, lemmas), but ommiting the proofs. In this work we
partially adhere to this vision to delineate our formalization
strategy:
Since the proofs that the  axioms hold in generic extensions
is independent of the \emph{proofs} of the fundamental theorems of
forcing, we assumed these results for the time being. Let us remark
that those theorems depend on the definition of a function $\forceisa$
from formulas to formulas which is, by itself, quite demanding; the
formalization of it and of the fundamental theorems of forcing % roughly
comprises barely less than a half of our full project.
